### 2020-02-14 CalNotes.tex
Chebyshev

In this section, we present how to correct the faulty behaviour highlighted by Theorem \ref{thm:BayesianBias} with filtering techniques. In particular, we employ the same strategy that led to the estimator $\widetilde A^\epl_k(T)$ presented in Section \ref{sec:FilterMLE_2}. Therefore, we modify the likelihood function as
\begin{equation}
	\widetilde p(X^\epl \mid A) = \exp\left(-\frac{\widetilde I(X^\epl\mid A)}{2\Sigma} \right), 
\end{equation}
where 
\begin{equation}
	\widetilde I(X^\epl \mid A) =\int_0^T A \cdot V'(Z^\epl_t) \dd X^\epl_t + \int_0^T \left(A \cdot V'(X^\epl_t)\right)^2 \dd t,
\end{equation}
so that the estimator $\widetilde A^\epl_k(T)$ introduced in \eqref{eq:AHatMixedTilde} is the MLE based on the new likelihood $\widetilde p(X^\epl \mid A)$. With this likelihood function, we obtain the modified posterior $\widetilde \mu_{T,\epl} = \mathcal N(\widetilde m_{T, \epl}, C_{T, \epl})$, whose parameters are given by
\begin{equation}
\begin{aligned}
	C_{T, \epl}^{-1} &= C_0^{-1} + T M, \\
	C_{T,\epl}^{-1}\widetilde m_{T,\epl} &= C_0^{-1}A_0 - T \widetilde h. 
\end{aligned}	
\end{equation}
Let us remark that the posterior $\widetilde \mu_{T,\epl}$ has the same covariance as $\mu_{T,\epl}$. The following result now holds.
\begin{theorem} Let the Assumptions of Theorem \ref{thm:mainTheoremTilde} hold. Then, the modified posterior measure $\widetilde\mu_{T,\epl} = \mathcal N(\widetilde m_{T,\epl}, C_{T,\epl})$ satisfies
	\begin{equation}
	\lim_{\epl \to 0}\lim_{T \to \infty} \widetilde \mu_{T,\epl}(\cdot) = \delta(\cdot - A), \quad \text{a.s.},
	\end{equation}
	weakly in the space of probability measures on $\R$, where $\delta$ is the Dirac measure on $\R$.
\end{theorem}
\begin{proof} The proof follows from the proof of Theorem \ref{thm:BayesianBias} and from Theorem \ref{thm:mainTheoremTilde}.
\end{proof}

@@

\begin{figure}[t]
	\centering
	\begin{tabular}{ccc}
		\includegraphics[]{Figures/KLNothing} & \includegraphics[]{Figures/KLFilt}  & \includegraphics[]{Figures/KLSubs}
	\end{tabular}
	\caption{Results for the four-dimensional parameters. From left to right the potential function estimated with the data itself, the filter, subsampled data.}
	\label{fig:KLStyle}
\end{figure}

We consider $N = 4$ and the potential function $V(x)$ as in \eqref{eq:Potential} with
\begin{equation}
	V_i(x) = \frac{x^{2i}}{2i}, \quad i =1, \ldots, 4,
\end{equation}
so that $\alpha \cdot V'(x) = \alpha_1 x + \alpha_2 x^3 + \alpha_3 x^5 + \alpha_4 x^7$, with $\alpha = (-2, -1, 1, 1/2)$. We consider $\epl = 0.05$, the fast potential $p(y) = \cos(y)$ and simulate a trajectory of $X^\epl$ for $0 \leq t \leq T$ with $T = 10^3$ employing the Euler--Maruyama method with time step $\Delta_t = \epl^3$. We compute the MLE of the drift coefficient $A \in \R^4$ of the homogenized equation \eqref{eq:SDE_HOM} employing 
\begin{enumerate}
	\item the data $X^\epl$ itself,
	\item subsampled data with subsampling parameter $\delta = \epl^{2/3}$,
	\item filtered data $Z^\epl$ computed with $\beta = 1$ and $\delta = 0.2$.
\end{enumerate}
In particular, we pick this specific value of $\delta$ for the subsampling following the optimality criterion given in \cite{PaS07}. Results, given in Figure \ref{fig:KLStyle}, show that the filter-based estimation captures well the homogenized potential.

>>>>

In this section, we present how to correct the faulty behaviour highlighted by Theorem \ref{thm:BayesianBias} with filtering techniques. In particular, we employ the same strategy that led to the estimator $\widetilde A^\epl_k(T)$ presented in Section \ref{sec:FilterMLE_2}. Therefore, we modify the likelihood function as
\begin{equation}
	\widetilde p(X^\epl \mid A) = \exp\left(-\frac{\widetilde I(X^\epl\mid A)}{2\Sigma} \right), 
\end{equation}
where 
\begin{equation}
	\widetilde I(X^\epl \mid A) =\int_0^T A \cdot V'(Z^\epl_t) \dd X^\epl_t + \int_0^T \left(A \cdot V'(X^\epl_t)\right)^2 \dd t,
\end{equation}
so that the estimator $\widetilde A^\epl_k(T)$ introduced in \eqref{eq:AHatMixedTilde} is the MLE based on the new likelihood $\widetilde p(X^\epl \mid A)$. With this likelihood function, we obtain the modified posterior $\widetilde \mu_{T,\epl} = \mathcal N(\widetilde m_{T, \epl}, C_{T, \epl})$, whose parameters are given by
\begin{equation}
\begin{aligned}
	C_{T, \epl}^{-1} &= C_0^{-1} + T M, \\
	C_{T,\epl}^{-1}\widetilde m_{T,\epl} &= C_0^{-1}A_0 - T \widetilde h. 
\end{aligned}	
\end{equation}
Let us remark that the posterior $\widetilde \mu_{T,\epl}$ has the same covariance as $\mu_{T,\epl}$. The following result now holds.
\begin{theorem} Let the Assumptions of Theorem \ref{thm:mainTheoremTilde} hold. Then, the modified posterior measure $\widetilde\mu_{T,\epl} = \mathcal N(\widetilde m_{T,\epl}, C_{T,\epl})$ satisfies
	\begin{equation}
	\lim_{\epl \to 0}\lim_{T \to \infty} \widetilde \mu_{T,\epl}(\cdot) = \delta(\cdot - A), \quad \text{a.s.},
	\end{equation}
	weakly in the space of probability measures on $\R^N$, where $\delta$ is the Dirac measure on $\R^N$.
\end{theorem}
\begin{proof} The proof follows from the proof of Theorem \ref{thm:BayesianBias} and from Theorem \ref{thm:mainTheoremTilde}.
\end{proof}

@@

\begin{figure}[t]
	\centering
	\begin{tabular}{ccc}
		\includegraphics[]{Figures/KLNothing} & \includegraphics[]{Figures/KLFilt}  & \includegraphics[]{Figures/KLSubs}
	\end{tabular}
	\caption{Results for the four-dimensional parameter. From left to right the potential function estimated with the data itself, the filter, subsampled data.}
	\label{fig:KLStyle}
\end{figure}

\begin{table}[t]
	\centering
	\begin{tabular}{ccccc}
		\toprule
		Coefficient & Truth & No preprocessing & Filtering & Subsampling  \\ 
		\midrule
		$A_1$ & -0.62 & -0.92 & -0.70 & -0.59\\
		$A_2$ & -0.31 & -0.70 & -0.27 & 0.05 \\
		$A_3$ & 0.31  & 0.55  & 0.31  & 0.14 \\
		$A_4$ & 0.62  & 1.22  & 0.57  & 0.13 \\
		\bottomrule
	\end{tabular}
	\caption{Numerical results for the four-dimensional parameter rounded to the second significant figure.}
	\label{tab:KLStyle}
\end{table}

Let us consider the Chebyshev polynomials of the first kind, i.e., the polynomials $T_i\colon \R \to \R$, $i=0, 1, \ldots$, defined by the recurrence relation
\begin{equation}
	T_0(x) = 1, \quad T_1(x) = x, \quad T_{i+1}(x) = 2xT_i(x) - T_{i-1}(x).
\end{equation}
We consider the potential function $V(x)$ as in \eqref{eq:Potential} with
\begin{equation}
	V_i(x) = T_i(x), \quad i =1, \ldots, 4.
\end{equation}
This potential function satisfies Assumption \ref{as:regularity} whenever $N$ is even and if the leading coefficient $\alpha_N$ is positive. We set $N = 4$ and the drift coefficient $\alpha = (-1, -1/2, 1/2, 1)$. With this drift coefficient, the potential function is of the bistable kind. Moreover, we set $\epl = 0.05$, the diffusion coefficient $\sigma = 1$, the fast potential $p(y) = \cos(y)$ and simulate a trajectory of $X^\epl$ for $0 \leq t \leq T$ with $T = 10^3$ employing the Euler--Maruyama method with time step $\Delta_t = \epl^3$. We estimate the drift coefficient $A \in \R^4$ with the estimators
\begin{enumerate}
	\item $\widehat A^\epl(T)$ based on the data $X^\epl$ itself,
	\item $\widehat A^\epl_\delta(T)$ based on subsampled data with subsampling parameter $\delta = \epl^{2/3}$,
	\item $\widehat A^\epl_k(T)$ based on filtered data $Z^\epl$ computed with $\beta = 1$ and $\delta = 1$.
\end{enumerate}
In particular, we pick this specific value of $\delta$ for the subsampling following the optimality criterion given in \cite{PaS07}. Results, given in Figure \ref{fig:KLStyle} and Table \ref{tab:KLStyle}, show that the filter-based estimation captures well the homogenized potential as well as the coefficient $A$. Moreover, it is possible to remark the negative result given by Theorem \ref{thm:Bias} holds in practice, i.e., with no pre-processing the estimator $\widehat A^\epl(T)$ tends to the drift coefficient $\alpha$ of the unhomogenized equation.
